### 2024-03-20 src/beam/tables/minimization_data.tex
add minimization tables

\documentclass{standalone}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.18}

% Setup siunitx:
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

\begin{document}

\pgfplotstabletypeset[
  col sep=comma,
  % columns={Model,Iterations,Evaluations,Time,{Output J}},
  display columns/3/.style={column name={Time}, precision=3},
  display columns/4/.style={column name={Output $J$}, sci zerofill, precision=3},
  every head row/.style={
    before row={\toprule}, % have a rule at top
    % set unit for each column
    after row={
        & - & - & \si{\second} & \si{\newton\milli\metre}\\
    \midrule}
    },
  every last row/.style={after row=\bottomrule}, % rule at bottom
  column type=r,
  columns/Model/.style={string type,column type=l},
]{./work/beam/minimization_data.csv}

\end{document}