### 2024-03-20 src/beam/tables/minimization_comparison.tex

\documentclass{standalone}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{pgfplotstable}
\usepackage{amsmath,physics}
\pgfplotsset{compat=1.18}

% Setup siunitx:
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 2, % to 2 places
}

\begin{document}

\pgfplotstabletypeset[
  % FIXME using column type {S} does not lead to desired result 
  % columns/{Beam example}/.style={
  %     column type={S},
  %     sci zerofill,
  %     precision=3,
  % },
  multicolumn names,
  col sep=comma,
  column type=r,
  % columns={Model,Iterations,Evaluations,Time,{Output J}},
  display columns/0/.style={string type, column name={}, column type=l},
  display columns/1/.style={string type, column name={}, column type=l},
  display columns/2/.style={column name={Beam example}, sci zerofill, precision=4},
  every head row/.style={
    before row={\toprule}, % have a rule at top
    after row={\midrule},
    },
  every last row/.style={after row=\bottomrule}, % rule at bottom
]{./work/beam/minimization_comparison.csv}

\end{document}